\section{Cumulative distribution Functions}

The functions related to this chapter are declared in \verb!pnl/pnl_cdf.h!.

For various distribution functions, we provide functions named
\var{pnl_cdf_xxx} where \var{xxx} is the abbreviation of the distribution
name. All these functions are based on the same prototype
\begin{equation*}
  p = 1-q; \quad p = \int^x density(u) du
\end{equation*}

\begin{itemize}
\item \var{which} If \var{which=1}, it computes \var{p} and \var{q}. If
  \var{which=2}, it computes \var{x}. For higher values of \var{which} it
  computes one the parameters characterizing the distribution using all the
  others, \var{p, q, x}.
\item \var{p} the probability $\int^x density(u) du $
\item \var{q} $= 1 - p$
\item \var{x} the upper bound of the integral
\item \var{status} an integer which indicates on exit the success of the
  computation. (0) if calculation completed correctly. (-I) if the input
  parameter number I was out of range. (1) if the answer appears to be lower
  than the lowest search bound.  (2) if the answer appears to be higher than
  the greatest search bound.  (3) if $p + q \ne 1$.
\item \var{bound} is undefined if STATUS is 0.  Bound exceeded by parameter
  number I if STATUS is negative. Lower search bound if STATUS is 1.  Upper
  search bound if STATUS is 2.
\end{itemize}

\begin{itemize}
\item \describefun{void}{pnl_cdf_bet}{int \ptr which, double \ptr p, double
    \ptr q, double \ptr x, double \ptr y, double \ptr a, double \ptr b,
    int \ptr status, double \ptr bound}
  \sshortdescribe Cumulative Distribution Function BETA distribution.

\item \describefun{void}{pnl_cdf_bin}{int \ptr which, double \ptr p, double
    \ptr q, double \ptr x, double \ptr xn, double \ptr pr, double
    \ptr ompr, int \ptr status, double \ptr bound}
  \sshortdescribe Cumulative Distribution Function BINa distribution.

\item \describefun{void}{pnl_cdf_chi}{int \ptr which, double \ptr p, double
    \ptr q, double \ptr x, double \ptr df, int \ptr status, double
    \ptr bound}
  \sshortdescribe Cumulative Distribution Function CHI-Square distribution.

\item \describefun{void}{pnl_cdf_chn}{int \ptr which, double \ptr p, double
    \ptr q, double \ptr x, double \ptr df, double \ptr pnonc, int
    \ptr status, double \ptr bound}
  \sshortdescribe Cumulative Distribution Function Non-central Chi-Square distribution.

\item \describefun{void}{pnl_cdf_f}{int \ptr which, double \ptr p, double
    \ptr q, double \ptr x, double \ptr dfn, double \ptr dfd, int
    \ptr status, double \ptr bound}
  \sshortdescribe Cumulative Distribution Function F distribution.

\item \describefun{void}{pnl_cdf_fnc}{int \ptr which, double \ptr p, double
    \ptr q, double \ptr x, double \ptr dfn, double \ptr dfd, double
    \ptr pnonc, int \ptr status, double \ptr bound}
  \sshortdescribe Cumulative Distribution Function Non-central F distribution.

\item \describefun{void}{pnl_cdf_gam}{int \ptr which, double \ptr p, double
    \ptr q, double \ptr x, double \ptr shape, double \ptr rate, int
    \ptr status, double \ptr bound}
  \sshortdescribe Cumulative Distribution Function GAMma distribution. Note that
  the parameter \var{rate} is \var{1/scale}. The density writes 
  $f(x)= 1/(s^a \Gamma(a)) x^{a-1} e^{-x/s}$ with \var{scale=s} and \var{shape=1/rate=a}.

\item \describefun{void}{pnl_cdf_nbn}{int \ptr which, double \ptr p, double
    \ptr q, double \ptr x, double \ptr xn, double \ptr pr, double
    \ptr ompr, int \ptr status, double \ptr bound}
  \sshortdescribe Cumulative Distribution Function Negative BiNomial distribution.

\item \describefun{void}{pnl_cdf_nor}{int \ptr which, double \ptr p, double
    \ptr q, double \ptr x, double \ptr mean, double \ptr sd, int
    \ptr status, double \ptr bound}
  \sshortdescribe Cumulative Distribution Function NORmal distribution.

\item \describefun{void}{pnl_cdf_poi}{int \ptr which, double \ptr p, double
    \ptr q, double \ptr x, double \ptr xlam, int \ptr status, double
    \ptr bound}
  \sshortdescribe Cumulative Distribution Function POIsson distribution.

\item \describefun{void}{pnl_cdf_t}{int \ptr which, double \ptr p, double
    \ptr q, double \ptr x, double \ptr df, int \ptr status, double
    \ptr bound}
  \sshortdescribe Cumulative Distribution Function T distribution.

\item \describefun{double}{pnl_cdfchi2n}{double x, double df, double ncparam}
  \sshortdescribe Compute the cumulative density function at \var{x} of the
  non central $\chi^2$ distribution with \var{df} degrees of freedom and non
  centrality parameter \var{ncparam}.

\item \describefun{void}{pnl_cdfbchi2n}{double x, double df, double ncparam, double
    beta, double \ptr P}
  \sshortdescribe Store in \var{P} the cumulative density function at \var{x}
  of the random variable \var{beta \ptr X} where \var{X} is non central $\chi^2$
  random variable with \var{df} degrees of freedom and non
  centrality parameter \var{ncparam}.


\item \describefun{double}{pnl_normal_density}{double x}
  \sshortdescribe Normal density function.

\item \describefun{double}{pnl_cdfnor}{double x}
  \sshortdescribe Cumulative normal distribution function.

\item \describefun{double}{pnl_cdf2nor}{double a, double b, double r}
  \sshortdescribe Cumulative bivariate normal distribution function, returns
  $\frac{1} {2\pi \sqrt{1-r^2}} \int_{- \infty}^a\int_{- \infty}^b e^{-
    \frac{x^2 - 2 r xy+y^2} {2(1-r^2)} } dxdy.$

\item \describefun{double}{pnl_inv_cdfnor}{double x}
  \sshortdescribe Inverse of the cumulative normal distribution function.
\end{itemize}

